---
title: "Describe and compare upper and lower lung parameters."
output: html_notebook
---



Load up packages.
```{r}
pacman::p_load(tidyr, dplyr, ggplot2, plyr, nlme, reshape2, FactoMineR, BSDA)
```

Get the data.
```{r}
raw_data <- read.csv2("C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/FD_data/Expi.csv")
raw_data2 <- read.csv("C:/Users/csuehs/Dropbox/SCANN'AIR_princeps/data-verification/Data_scannair_NA_corrections_isaetcarey_20170427.csv")


```

Extract the appropriate variables. We are interested in raw FD values for attentuation at -850HU and according to two factors: 

* inspiratory versus expiratory measures;
* upper versus lower lung.
      ---slice 1
      ---slice 10
```{r}

df_fd <- data.frame(raw_data$numpat,
                  raw_data$I.850.H1,
                  raw_data$I.850.H10,
                  raw_data$I.850.L1,
                  raw_data$I.850.L10,
                  raw_data$E.850.H1,
                  raw_data$E.850.H10,
                  raw_data$E.850.L1,
                  raw_data$E.850.L10)

colnames(df_fd) <- c("patient",
                   "fd_i_up_1", 
                    "fd_i_up_10", 
                    "fd_i_lo_1", 
                    "fd_i_lo_10", 
                    "fd_e_up_1", 
                    "fd_e_up_10", 
                    "fd_e_lo_1", 
                    "fd_e_lo_10")



df_fd$fd_ei_up_1 <- df_fd$fd_e_up_1 / df_fd$fd_i_up_1
df_fd$fd_ei_up_10 <- df_fd$fd_e_up_10 / df_fd$fd_i_up_10

df_fd$fd_ei_lo_1 <- df_fd$fd_e_lo_1 / df_fd$fd_i_lo_1
df_fd$fd_ei_lo_10 <- df_fd$fd_e_lo_10 / df_fd$fd_i_lo_10


df_fd$fd_i_up <- (df_fd$fd_i_up_1 + df_fd$fd_i_up_10)/2
df_fd$fd_i_lo <- (df_fd$fd_i_lo_1 + df_fd$fd_i_lo_10)/2

df_fd$fd_e_up <- (df_fd$fd_e_up_1 + df_fd$fd_e_up_10)/2
df_fd$fd_e_lo <- (df_fd$fd_e_lo_1 + df_fd$fd_e_lo_10)/2

df_fd$fd_ei_up <- (df_fd$fd_ei_up_1+df_fd$fd_ei_up_10)/2
df_fd$fd_ei_lo <- (df_fd$fd_ei_lo_1+df_fd$fd_ei_lo_10)/2

df_fd$fd_i_delta <- df_fd$fd_i_up - df_fd$fd_i_lo  
df_fd$fd_e_delta <- df_fd$fd_e_up - df_fd$fd_e_lo 
df_fd$fd_ei_delta <- df_fd$fd_ei_up - df_fd$fd_ei_lo 

df_fd$fd_i_av <- (df_fd$fd_i_up + df_fd$fd_i_lo) /2
df_fd$fd_e_av <- (df_fd$fd_e_up + df_fd$fd_e_lo) /2
df_fd$fd_ei_av <- (df_fd$fd_ei_up + df_fd$fd_ei_lo) /2

df_fd$fd_i_bias <- (df_fd$fd_i_delta / df_fd$fd_i_av) *100
df_fd$fd_e_bias <- (df_fd$fd_e_delta / df_fd$fd_e_av) *100
df_fd$fd_ei_bias <- (df_fd$fd_ei_delta / df_fd$fd_ei_av) *100

df_fd[,grep("_1", names(df_fd))] <- NULL
df_fd[,grep("delta", names(df_fd))] <- NULL
df_fd[,grep("av2", names(df_fd))] <- NULL


head(df_fd)

```

Extract the appropriate MLD variables.
```{r}
df_mld <- data.frame(raw_data2$numpat,
                     raw_data2$I_MLD_H_1,
                     raw_data2$I_MLD_B_1,
                     raw_data2$E_MLD_H_1,
                     raw_data2$E_MLD_B_1)
    

colnames(df_mld) <- c("patient",
                      "mld_i_up",
                      "mld_i_lo",
                      "mld_e_up",
                      "mld_e_lo")

df_mld$mld_ei_up <- df_mld$mld_e_up/df_mld$mld_i_up
df_mld$mld_ei_lo <- df_mld$mld_e_lo/df_mld$mld_i_lo

df_mld$mld_i_delta <- df_mld$mld_i_up - df_mld$mld_i_lo 
df_mld$mld_e_delta <- df_mld$mld_e_up - df_mld$mld_e_lo 
df_mld$mld_ei_delta <- df_mld$mld_ei_up - df_mld$mld_ei_lo 

df_mld$mld_i_av <- (df_mld$mld_i_up + df_mld$mld_i_lo)/2 
df_mld$mld_e_av <- (df_mld$mld_e_up + df_mld$mld_e_lo)/2
df_mld$mld_ei_av <- (df_mld$mld_ei_up + df_mld$mld_ei_lo)/2

df_mld$mld_i_bias <- (df_mld$mld_i_delta / df_mld$mld_i_av) *100
df_mld$mld_e_bias <- (df_mld$mld_e_delta / df_mld$mld_e_av) *100
df_mld$mld_ei_bias <- (df_mld$mld_ei_delta / df_mld$mld_ei_av) *100

df_mld[,grep("delta", names(df_mld))] <- NULL

head(df_mld)

```

Describe the centrality and distribution of fd measures.
```{r}
temp <- df_fd
temp$patient <- NULL

eff <- function(x) {length(which(!is.na(x)))}
quartile_1 <- function(x) {quantile(x, probs = seq(0,1,0.25), na.rm = TRUE)[2]}
quartile_3 <- function(x) {quantile(x, probs = seq(0,1,0.25), na.rm = TRUE)[4]}
shap <- function(x) {as.numeric(shapiro.test(x)[2])}

count<- sapply(temp, eff)
minimum <- sapply(temp, min, na.rm = TRUE)
maximum <- sapply(temp, max, na.rm = TRUE)
mean <- sapply(temp, mean, na.rm = TRUE)
st_d <- sapply(temp, sd, na.rm = TRUE)
median <- sapply(temp, median, na.rm = TRUE)
quartile_25p <- sapply(temp, quartile_1)
quartile_75p <- sapply(temp, quartile_3)
shapiro_p <- sapply(temp, shap)

options(scipen = 999)
descriptive_total_pop <- data.frame(count, minimum, maximum, mean, 
                     st_d, median, quartile_25p, quartile_75p, shapiro_p)

rm(count, minimum, maximum, mean, st_d, median, quartile_25p, quartile_75p, shapiro_p)

write.csv(descriptive_total_pop, file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/stats-tables/fd_descriptive_data_total_population.csv")

write.csv2(descriptive_total_pop, file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/stats-tables/fd_descriptive_data_total_population_csv2.csv")

rm(temp)

descriptive_total_pop
```

Describe the centrality and distribution of mld measures.
```{r}
temp <- df_mld
temp$patient <- NULL

count<- sapply(temp, eff)
minimum <- sapply(temp, min, na.rm = TRUE)
maximum <- sapply(temp, max, na.rm = TRUE)
mean <- sapply(temp, mean, na.rm = TRUE)
st_d <- sapply(temp, sd, na.rm = TRUE)
median <- sapply(temp, median, na.rm = TRUE)
quartile_25p <- sapply(temp, quartile_1)
quartile_75p <- sapply(temp, quartile_3)
shapiro_p <- sapply(temp, shap)

options(scipen = 999)
descriptive_total_pop <- data.frame(count, minimum, maximum, mean, 
                     st_d, median, quartile_25p, quartile_75p, shapiro_p)

rm(count, minimum, maximum, mean, st_d, median, quartile_25p, quartile_75p, shapiro_p)

write.csv(descriptive_total_pop, file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/stats-tables/mld_descriptive_data_total_population.csv")

write.csv2(descriptive_total_pop, file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/stats-tables/mld_descriptive_data_total_population_csv2.csv")

rm(temp)

descriptive_total_pop
```


Compare upper and lower lungs (paired tests)

```{r}

t.test(x = df_mld$mld_i_up, y = df_mld$mld_i_lo, paired = TRUE)
t.test(x = df_mld$mld_e_up, y = df_mld$mld_e_lo, paired = TRUE)
t.test(x = df_mld$mld_ei_up, y = df_mld$mld_ei_lo, paired = TRUE)


wilcox.test(x = df_fd$fd_i_up, y = df_fd$fd_i_lo, paired = TRUE)
wilcox.test(x = df_fd$fd_e_up, y = df_fd$fd_e_lo, paired = TRUE)
wilcox.test(x = df_fd$fd_ei_up, y = df_fd$fd_ei_lo, paired = TRUE)
```




Get the proportions of zeros in variables.
```{r}

trap_i_up <- df_fd$fd_i_up
trap_i_up[trap_i_up > 0] <- "yes"
trap_i_up[trap_i_up == 0] <- "no"

trap_e_up <- df_fd$fd_e_up
trap_e_up[trap_e_up > 0] <- "yes"
trap_e_up[trap_e_up == 0] <- "no"

trap_i_lo <- df_fd$fd_i_lo
trap_i_lo[trap_i_lo > 0] <- "yes"
trap_i_lo[trap_i_lo == 0] <- "no"

trap_e_lo <- df_fd$fd_e_lo
trap_e_lo[trap_e_lo > 0] <- "yes"
trap_e_lo[trap_e_lo == 0] <- "no"

```

Upper versus lower tests for proportion airtrapping detection.

```{r}

mcnemar.test(x = trap_e_up, y = trap_e_lo)

```































